% Paste-ready version aligned to the current manuscript flow.
% Notes:
% - This version avoids repeatedly using the phrase "wireless SFL".
% - Instead, it describes a standard SFL framework in which clients
%   communicate with the server over wireless uplink channels.
% - If your figure labels differ, replace fig:system_model and
%   fig:proposed_encoder accordingly.

% ------------------------------------------------------------------
% Fig. 1 caption
% ------------------------------------------------------------------
% \caption{Generic split federated learning framework considered in this work. Multiple clients collaboratively train a split model by transmitting smashed data to the server over wireless uplink channels and receiving intermediate gradients for backpropagation.}
% \label{fig:system_model}

% ------------------------------------------------------------------
% Fig. 3 caption
% ------------------------------------------------------------------
% \caption{Proposed CA-SSFL semantic encoder with VIB-driven semantic masking, SNR-driven channel masking, and dynamic slicing.}
% \label{fig:proposed_encoder}

% ------------------------------------------------------------------
% Section II
% ------------------------------------------------------------------
\section{System Model}

We consider an SFL framework that combines split learning and federated aggregation \cite{gupta2018distributed,thapa2022splitfed}, as illustrated in Fig.~\ref{fig:system_model}. Let $\mathcal{I}=\{1,\ldots,I\}$ denote the client set, and let $\mathcal{D}_i$ denote the local dataset at client $i \in \mathcal{I}$. In the considered setting, clients communicate with the server over wireless uplink channels. The global model is partitioned into a client-side sub-model $\phi_c$ and a server-side sub-model $\phi_s$.

For an input sample $x$, the client first computes an intermediate feature representation, commonly referred to as smashed data, as
\begin{equation}
f = \phi_c(x).
\end{equation}
The smashed data $f$ is then transmitted to the server through the wireless uplink. During transmission, the received representation at the server is corrupted by channel noise and fading, and is modeled as
\begin{equation}
f' = h \cdot f + n,
\end{equation}
where $n \sim \mathcal{CN}(0,\sigma_n^2 I)$ denotes additive white Gaussian noise and $h$ denotes the channel coefficient. In particular, $h=1$ for the AWGN channel and $h \sim \mathcal{CN}(0,1)$ for the Rayleigh fading channel. Based on the received representation, the server performs the remaining layers of the split model to produce the final prediction
\begin{equation}
\hat{y} = \phi_s(f').
\end{equation}

This split architecture reduces the computational burden on client devices by offloading most of the deep neural network computation to the server, which makes it attractive for resource-limited IoT environments. In the generic SFL setting, the split model is trained end-to-end using a standard supervised objective defined on the server output. For a ground-truth label $y$, the task loss is written as
\begin{equation}
\mathcal{L}_{\mathrm{task}} = \ell(\hat{y}, y),
\end{equation}
where $\ell(\cdot,\cdot)$ denotes a classification loss such as cross-entropy. During training, the server computes the loss from the final prediction and propagates the intermediate gradient back through the split point so that both $\phi_s$ and $\phi_c$ are jointly optimized. After local updates, the client-side parameters are periodically aggregated across participating clients in a federated manner to form the global model for the next communication round.

Although SFL is effective in reducing on-device computation, its communication cost remains substantial because high-dimensional smashed data must be transmitted at every training iteration \cite{gao2020end}. This problem becomes more severe when the client-server exchange takes place over wireless channels, where limited bandwidth and channel noise can significantly degrade transmission efficiency and downstream task performance. These challenges motivate the proposed channel-aware semantic transmission mechanism described in the next section.

% ------------------------------------------------------------------
% Section III
% ------------------------------------------------------------------
\section{Proposed Method}

\subsection{Proposed CA-SSFL Architecture}
Building on the SFL framework in Section II, we propose a channel-aware semantic split federated learning (CA-SSFL) architecture to reduce communication overhead while preserving task-relevant information under dynamic channel conditions. Motivated by task-oriented semantic communication \cite{qin2021semantic,gunduz2022beyond}, the proposed method augments the conventional client-server split model with a semantic encoder at the client side and a semantic decoder at the server side, as illustrated in Fig.~\ref{fig:proposed_encoder}. Accordingly, the client-side model is defined as $\Theta_C = \{\phi_c, \theta_{en}\}$, where $\phi_c$ extracts smashed data and $\theta_{en}$ converts it into a transmission-efficient latent representation. The server-side model is defined as $\Theta_S = \{\theta_{de}, \phi_s\}$, where $\theta_{de}$ reconstructs the received latent representation and $\phi_s$ performs the downstream prediction task. Therefore, the overall model is given by
\begin{equation}
\Theta = \{\phi_c, \theta_{en}, \theta_{de}, \phi_s\}.
\end{equation}

Unlike conventional SFL, the proposed CA-SSFL incorporates two key mechanisms in the client-side semantic encoder: 1) a VIB-based semantic masking module that estimates the task relevance of latent features, and 2) an SNR-driven channel masking module that determines their transmission feasibility under the current wireless condition. The final transmission support is determined by combining these two masks as
\begin{equation}
M_{\mathrm{final}} = M_{\mathrm{vib}} \odot M_{\mathrm{ch}},
\end{equation}
so that only the feature dimensions that are both semantically important and channel-feasible are selected for transmission. The selected latent features are then sent through a dynamic slicing operation, and the server restores the sparse representation to its original dimensionality before processing it through the semantic decoder and the remaining server-side network. The detailed operations of the semantic encoder, channel masking, dynamic slicing, and end-to-end training objective are presented in the following subsections.

\subsection{VIB-Driven Semantic Masking Module}
As illustrated in Fig.~\ref{fig:proposed_encoder}, the semantic encoder first transforms the flattened smashed data $f_{\mathrm{flt}} \in \mathbb{R}^{K}$ into the latent statistics of a variational representation, namely the mean $\mu$ and variance $\sigma^2$. To enable stochastic training, we adopt the reparameterization trick \cite{kingma2013auto} and sample the latent vector as
\begin{equation}
z = \mu + \sigma \odot \epsilon, \quad \epsilon \sim \mathcal{N}(0, I).
\end{equation}

Rather than treating all latent dimensions equally, CA-SSFL evaluates their semantic importance through dimension-wise KL divergence between the posterior and a standard normal prior following the variational information bottleneck principle \cite{alemi2017deep}. The KL divergence for the $k$-th dimension is given by
\begin{equation}
D_{\mathrm{KL}}^{(k)} = -\frac{1}{2}\left(1 + \log \sigma_k^2 - \mu_k^2 - \sigma_k^2\right).
\end{equation}
In implementation, the latent statistics used for KL evaluation are modulated by the current channel condition through the channel-aware encoder path. Therefore, the semantic importance is computed from channel-modulated latent statistics rather than from a strictly channel-independent branch. Based on this score, dimensions with KL values larger than a predefined semantic threshold $\tau_{\mathrm{VIB}}$ are retained, yielding the binary semantic mask
\begin{equation}
M_{\mathrm{vib}}^{(k)} = \mathbb{I}\left(D_{\mathrm{KL}}^{(k)} > \tau_{\mathrm{VIB}}\right).
\end{equation}

\subsection{SNR-Driven Channel Masking Module}
To reflect the instantaneous wireless condition, the normalized SNR $s_{\mathrm{norm}} \in [0,1]$ is fed into a lightweight masking path that generates a channel-feasibility score for each latent dimension. Specifically, this path produces a probability vector $P \in [0,1]^K$, where each element indicates how likely the corresponding latent dimension is to be reliably transmitted under the current channel state.

To adapt the transmission budget, we introduce a dynamic threshold
\begin{equation}
\tau_{\mathrm{ch}} = \tau_{\max} - s_{\mathrm{norm}}(\tau_{\max}-\tau_{\min}).
\end{equation}
Under low-SNR conditions, $\tau_{\mathrm{ch}}$ approaches $\tau_{\max}$, which enforces stronger compression. Under high-SNR conditions, it approaches $\tau_{\min}$, allowing more semantic features to be transmitted. The binary channel mask is then obtained as
\begin{equation}
M_{\mathrm{ch}}^{(k)} = \mathbb{I}\left(P^{(k)} > \tau_{\mathrm{ch}}\right),
\end{equation}
so that this module serves as a channel-dependent feasibility gate.

\subsection{Dual-Masking and Dynamic Slicing}
To jointly account for semantic relevance and transmission feasibility, the final transmission support is determined by the element-wise product of the semantic mask and the channel mask:
\begin{equation}
M_{\mathrm{final}} = M_{\mathrm{vib}} \odot M_{\mathrm{ch}}.
\end{equation}
The sampled latent vector is then masked as
\begin{equation}
z_{\mathrm{masked}} = z \odot M_{\mathrm{final}}.
\end{equation}
To realize actual communication reduction, the client transmits only the active latent values and their indices instead of sending the dense masked vector. At the server, the sparse latent representation is restored to the original dimensionality through zero-filling before being processed by the semantic decoder.

Since extreme sparsity can weaken the backpropagated learning signal, we additionally apply sparsity-aware gradient compensation. Let $p$ denote the active feature ratio. During backpropagation, the restored gradient is scaled by
\begin{equation}
\frac{1}{\max(p, 0.01)},
\end{equation}
so that the effective gradient magnitude remains stable even when the transmission budget varies significantly, following the intuition of inverted dropout scaling \cite{gradientScale}.

\subsection{Loss Function Design}
The end-to-end training objective of CA-SSFL combines the downstream task loss with a VIB-based compression penalty. Specifically, we minimize
\begin{equation}
\mathcal{L}_{\mathrm{total}} = \mathcal{L}_{\mathrm{task}} + \beta \mathcal{L}_{\mathrm{comp}},
\end{equation}
where $\beta$ controls the trade-off between task performance and information compression. Following the variational information bottleneck formulation \cite{alemi2017deep,wang2023performance}, the loss can be written as
\begin{equation}
\mathcal{L}_{\mathrm{total}}
=
\mathbb{E}_{z \sim p_{\theta_{en}}(z|\phi_c(x))}
\left[-\log q_{\theta_{de},\phi_s}(y|\tilde{z}')\right]
+ \beta D_{\mathrm{KL}}\!\left(p_{\theta_{en}}(z|\phi_c(x)) \,\|\, r(z)\right),
\end{equation}
where $\tilde{z}'$ denotes the channel-distorted transmitted latent representation and $r(z)$ is the standard normal prior. The first term encourages accurate downstream prediction after semantic transmission through the noisy wireless channel, while the second term penalizes redundant latent information through KL regularization. Unlike the generic SFL objective in Section II, this proposed objective explicitly learns transmission-efficient latent features jointly with semantic masking, channel masking, and semantic decoding. As a result, the encoder adaptively allocates the transmitted support according to both semantic relevance and instantaneous channel conditions.
